---
title: "Air pollution"
author: "Julian Flowers"
date: "24/02/2017"
output: 
  html_document: 
    number_sections: yes
    toc: yes
    toc_float: yes
---

```{r}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE, message = FALSE)
```

# Introduction

This short note illustrates three techniques:

* Batch downloads of data from the web using a *web scraping* approach.
* Organising the data into a tidy data frame.
* Using the `trelliscopejs` package to create a web page which contains a trend chart of pm25 concentrations for every local authority

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(trelliscopejs)
library(knitr)
library(plotly)
```

## *For* loop to download files for local authorities

DEFRA publish annual concentrations of pollutants for local authorities in the UK. Because the URLs of the each of the annual files have the same pattern we can write a *for loop* which downloads successive files, converts them into a tidy format, stacks the data and stores them in a data frame.


We can use the `DT` package to create an interactive table embedded in document which stores the data and is sortable and searchable.

```{r}
df <- data.frame()

for(year in 2010:2015){
  
  
ap <- read_csv(paste0("https://uk-air.defra.gov.uk/datastore/pcm/popwmpm25", year,"byUKlocalauthority.csv"), skip = 2)
ap <- ap %>% mutate(year = year) %>% select(-`LA code`)
ap <- ap %>% tidyr::gather(indicator, value, 1:3)
df <- bind_rows(df, ap)

}

df <- df %>%
  janitor::clean_names() %>%
  arrange(local_authority, indicator, year)

df %>%
  DT::datatable()
```

## Count number of values per year

```{r}

df %>%
  group_by(year) %>%
  count()
```

## Create faceted plot

There is a lot going on here:

1. In the second line we group the data by area, year and indicator
2. We calculate the mean pm25 values
3. We split the pm25 column into 3 parts
4. And drop the year (we already have a year column)
5. Then exclude the total values - just plot the non-anthropogenic and anthropogenic pm25 values
6. Then create a trend chart of pm25 values over time
7. Create a chart for each local authority and save as a web page - this uses the `facet_trelliscope` function - be patient it has to create over 400 charts.


```{r, lazy = TRUE }
df %>%
  group_by(local_authority, year, indicator) %>% 
  summarise(meanvals = mean(value, na.rm = TRUE)) %>%
  tidyr::separate(indicator, c("pm", "year1", "type"), sep = " ") %>%
  select(-year1) %>% 
  filter(type != "(total)" & year >2010) %>%
  ggplot(aes(year, meanvals)) + 
  geom_line(aes(colour = type))+
  geom_smooth(se = FALSE) +
  facet_trelliscope(~local_authority, self_contained = TRUE)
```
  






