---
title: "Diagnostic atlas of variation"
output:
  html_notebook: default
  html_document: default
editor_options:
  chunk_output_type: inline
---

# Introduction

Manipulating excel data in R can be difficult because often excel spreadsheets combine data storage, presentation and analysis in the same sheets and can be heavly formattted (merged columns, multiple column headers, multiple tables per sheet). The Atlas of Variation is an exampleof 'messy data'.
 
* Messy data - multiple sheets, variable formatting, variable field names


The objective is to extract CCG and other data to:
    + Put in PHOLIO format
    + New anlaysis

To do this we need to manipulate the data into tidy format.

Now new `R` tools for manipulating excel sheets - see <https://www.r-bloggers.com/how-to-use-jailbreakr/> which we will use to reorganise the spreadsheets and prepare the data for analysis.

## Data wrangling

Start by loading libraries.

```{r setup, include=FALSE}

knitr::opts_chunk$set(cache = TRUE, echo = TRUE, warning = FALSE)
options(scipen = 10, digits = 3)

library(readxl)
library(tidyverse)
if(!require(c("hadley/xml2","rsheets/linen",
                           "rsheets/cellranger",
                           "rsheets/rexcel",
                           "rsheets/jailbreakr")))devtools::install_github(c("hadley/xml2",
                           "rsheets/linen",
                           "rsheets/cellranger",
                           "rsheets/rexcel",
                           "rsheets/jailbreakr"))
library(rexcel)
library(jailbreakr)


```

and download the data from the [Fingertips website](https://fingertips.phe.org.uk/profile/atlas-of-variation).

```{r download}
url <- "https://fingertips.phe.org.uk/documents/DiagAtlasData_291116.xlsx"
destfile <- "aov_diag.xlsx"
aov_diag <- download.file(url, destfile)

```

Next we can look to see how many sheets are in the dataset - we'll use `rexcel_read_workbook` to read in the whole workbook - this takes a little while. W can see there are 38 sheets in the workbook. We have saved this to and `R` object we have called `workbook`.

```{r view sheets, message=FALSE, warning=FALSE}

workbook <- rexcel_read_workbook("aov_diag.xlsx", progress = FALSE)

sheet_names <- names(workbook$sheets) ## these are the sheet names

```

Let's look at sheet 1 to see what the structure is. The `jailbreakr` package has a `split_sheet` function, which extracts tables from the sheet.

```{r split sheet 1}
tables <- split_sheet(workbook$sheets$`1`)

glimpse(tables)


```

This identifies 4 subtables (look at `dim`) - one with 2 rows and 2 columns and three with 1050 rows and 15 columns. To extract the data values we need to do a bit more work. We can see the first table contains the title, and subsequent tables appear to have the same data structure but relate to different time periods. In the original spreadsheet these are separated by blank rows.

```{r extract values}

data_list <- map(tables, function(x) x$values())

## extract column names from 2nd table

col_names <- (data_list[[2]][1,]) %>% unlist

## join datasets together as a new table
map1 <- map_df(data_list[2:4], as.tibble)

## attach column names
colnames(map1) <- col_names


map1 <- map1 %>%
  slice(2:3151) %>% ## exclude header row
  unnest()

## get sheet title

title <- unlist(data_list[[1]][2,2])


title %>% unlist

## Attach to sheet

map1 <- map1 %>% 
  mutate(map = title) %>%
  select(Period = Year, `CCG code`, `CCG name`, Rate:map)

```

#### Simple analysis
Now we have a data frame for the first map which we can further analyse...

For example plotting data over time. We have quarterly and full year data - lets just plot the quarterly data.

This suggests that:

1. The rate of CT scanning has increased quarter by quarter
2. There was an extreme outlier earlier in the time series (Stafford and Surrounds CCG)
3. Judged by IQR variation in CT scanning rates between CCGs has increased over time series


```{r, fig.cap= "Boxplot of map1 over time"}
library(govstyle)


## which CCG has highest rate?
map1 %>%
  filter(stringr::str_detect(Period, "Q" )) %>%   ## identifies quarterly data
  group_by(Period) %>%
  top_n(wt = Rate, 1)


map1 %>%
  filter(stringr::str_detect(Period, "Q" )) %>%   ## identifies quarterly data
  ggplot(aes(Period, Rate)) +
  geom_boxplot(fill = "#006435") +
  theme_gov() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title =stringr::str_wrap(paste0(title1), 50))



```


## Scaling up

Now we have a method for extracting the data we can (try) to extract each table and combine the CCG data so we can look at the interrelationships between diagnostic tests. [I am doing this sheet by sheet but could be done programmatically altough there are several differnt patterns].

```{r cache=TRUE}

## Map 1
  tables <- split_sheet(workbook$sheets$`1`)
  data_set <- map(tables, function(x) x$values())
  col_names <- (data_set[[2]][1,]) %>% unlist
  map1 <- map_df(data_set[2:4], as.tibble)
  colnames(map1) <- col_names
map1 <- map1 %>%
    slice(2:3151) %>% ## exclude header row
    unnest()
  title1 <- unlist(data_set[[1]][2,2])
  title1 %>% unlist
map1 <- map1 %>% 
  mutate(map = title1) %>%
  select(Period, `CCG code`, `CCG name`, Rate:map)

## Map 2 
  tables <- split_sheet(workbook$sheets$`2`)
  data_set <- map(tables, function(x) x$values())
  col_names <- (data_set[[2]][1,]) %>% unlist
  map2 <- map_df(data_set[2:4], as.tibble)
  colnames(map2) <- col_names
map2 <- map2 %>%
    slice(2:3151) %>% ## exclude header row
    unnest()
  title1 <- unlist(data_set[[1]][2,2])
  title1 %>% unlist
map2 <- map2 %>% 
  mutate(map = title1) %>%
  select(Period, `CCG code`, `CCG name`, Rate:map)

## Map 3
tables <- split_sheet(workbook$sheets$`3`)
  data_set <- map(tables, function(x) x$values())
  col_names <- (data_set[[2]][1,]) %>% unlist
  map3 <- map_df(data_set[2:4], as.tibble)
  colnames(map3) <- col_names
map3 <- map3 %>%
    slice(2:3151) %>% ## exclude header row
    unnest()
  title1 <- unlist(data_set[[1]][2,2])
  title1 %>% unlist
map3 <- map3 %>% 
  mutate(map = title1) %>%
  select(Period, `CCG code`, `CCG name`, Rate:map)

## Map 16 - this differs - only 2 subtables
tables <- split_sheet(workbook$sheets$`16`)
glimpse(tables)
  data_set <- map(tables, function(x) x$values())
  col_names <- (data_set[[1]][3,]) %>% unlist
  map16 <- map_df(data_set[1], as.data.frame)
  colnames(map16) <- col_names
map16 <- map16 %>%
    slice(5:2103) %>%
    unnest()
  title1 <- unlist(data_set[[1]][2,2])
  title1 %>% unlist
map16 <- map16 %>% 
  mutate(map = title1) %>%
  select(Period = Year, `CCG code`, `CCG name`, `Observed procedures`:map)

## Map 17 - this differs - only 2 subtables
tables <- split_sheet(workbook$sheets$`17`)
glimpse(tables)
  data_set <- map(tables, function(x) x$values())
  col_names <- (data_set[[2]][1,]) %>% unlist
  map17 <- map_df(data_set[2], as.data.frame)
  colnames(map17) <- col_names
map17 <- map17 %>%
    slice(2:421) %>%
    unnest()
  title1 <- unlist(data_set[[1]][2,2])
  title1 %>% unlist
map17 <- map17 %>% 
  mutate(map = title1) %>%
  select(Period = Year, `CCG code`, `CCG name`, Rate:map)


## Map 19 - this differs - only 2 subtables
tables <- split_sheet(workbook$sheets$`19`)
glimpse(tables)
  data_set <- map(tables, function(x) x$values())
  col_names <- (data_set[[1]][3,]) %>% unlist
  map19 <- map_df(data_set[1], as.data.frame)
  colnames(map19) <- col_names
map19 <- map19 %>%
    slice(4:2093) %>%
    unnest()
  title1 <- unlist(data_set[[1]][2,2])
  title1 %>% unlist
map19 <- map19 %>% 
  mutate(map = title1, Year = as.character(Year)) %>%
  select(Period = Year, `CCG code`, `CCG name`, Rate = ISR, `95% lower`:map)

## Map 20 - this differs - only 2 subtables
tables <- split_sheet(workbook$sheets$`20`)
glimpse(tables)
  data_set <- map(tables, function(x) x$values())
  col_names <- (data_set[[2]][1,]) %>% unlist
  map20 <- map_df(data_set[2], as.data.frame)
  colnames(map20) <- col_names
map20 <- map20 %>%
    slice(2:2101) %>%
    unnest()
  title1 <- unlist(data_set[[1]][2,2])
  title1 %>% unlist
map20 <- map20 %>% 
  mutate(map = title1) %>%
  select(Period = Year, `CCG code`, `CCG name`, Rate = Percent, `95% lower`:map)

## Map 22 - this differs - only 2 subtables
tables <- split_sheet(workbook$sheets$`22`)
glimpse(tables)
  data_set <- map(tables, function(x) x$values())
  col_names <- (data_set[[1]][3,]) %>% unlist
  map22 <- map_df(data_set[1], as.data.frame)
  colnames(map22) <- col_names
map22 <- map22 %>%
    slice(4:2093) %>%
    unnest()
  title1 <- unlist(data_set[[1]][2,2])
  title1 %>% unlist
map22 <- map22 %>% 
  mutate(map = title1, Year = as.character(Year)) %>%
  select(Period = Year, `CCG code`, `CCG name`, Rate = ISR, `95% lower`:map)


## Map 23 - this differs - only 2 subtables
tables <- split_sheet(workbook$sheets$`23`)
glimpse(tables)
  data_set <- map(tables, function(x) x$values())
  col_names <- (data_set[[1]][3,]) %>% unlist
  map23 <- map_df(data_set[1], as.data.frame)
  colnames(map23) <- col_names
map23 <- map23 %>%
    slice(4:1683) %>%
    unnest()
  title1 <- unlist(data_set[[1]][2,2])
  title1 %>% unlist
map23 <- map23 %>% 
  mutate(map = title1) %>%
  select(Period = Year, `CCG code`, `CCG name`, Rate = DSR, `95% lower`:map)

## Map 24 - this differs - only 2 subtables
tables <- split_sheet(workbook$sheets$`24`)
glimpse(tables)
  data_set <- map(tables, function(x) x$values())
  col_names <- (data_set[[2]][1,]) %>% unlist
  map24 <- map_df(data_set[2:4], as.data.frame)
  colnames(map24) <- col_names
map24 <- map24 %>%
    slice(2:3151) %>%
    unnest()
  title1 <- unlist(data_set[[1]][2,2])
  title1 %>% unlist
map24 <- map24 %>% 
  mutate(map = title1) %>%
  select(Period = Year, `CCG code`, `CCG name`, Rate:map)


## Map 25 - this differs - only 2 subtables
tables <- split_sheet(workbook$sheets$`25`)
glimpse(tables)
  data_set <- map(tables, function(x) x$values())
  col_names <- (data_set[[2]][1,]) %>% unlist
  map25 <- map_df(data_set[2:4], as.data.frame)
  colnames(map25) <- col_names
map25 <- map25 %>%
    slice(2:3151) %>%
    unnest()
  title1 <- unlist(data_set[[1]][2,2])
  title1 %>% unlist
map25 <- map25 %>% 
  mutate(map = title1) %>%
  select(Period = Year, `CCG code`, `CCG name`, Rate:map)

## Map 26 - this differs - only 2 subtables
tables <- split_sheet(workbook$sheets$`26`)
glimpse(tables)
  data_set <- map(tables, function(x) x$values())
  col_names <- (data_set[[2]][1,]) %>% unlist
  map26 <- map_df(data_set[2], as.data.frame)
  colnames(map26) <- col_names
map26 <- map26 %>%
    slice(2:631) %>%
    unnest()
  title1 <- unlist(data_set[[1]][2,2])
  title1 %>% unlist
map26 <- map26 %>% 
  mutate(map = title1) %>%
  select(Period = Year, `CCG code`, `CCG name`, Rate = Percent, `95% lower`:map)

## Map 27 - this differs - only 2 subtables
tables <- split_sheet(workbook$sheets$`27`)
glimpse(tables)
  data_set <- map(tables, function(x) x$values())
  col_names <- (data_set[[2]][1,]) %>% unlist
  map27 <- map_df(data_set[2:4], as.data.frame)
  colnames(map27) <- col_names
map27 <- map27 %>%
    slice(2:3151) %>%
    unnest()
  title1 <- unlist(data_set[[1]][2,2])
  title1 %>% unlist
map27 <- map27 %>% 
  mutate(map = title1) %>%
  select(Period, `CCG code`, `CCG name`, Rate:map)

## Map 28 - this differs - only 2 subtables
tables <- split_sheet(workbook$sheets$`28`)
glimpse(tables)
  data_set <- map(tables, function(x) x$values())
  col_names <- (data_set[[2]][1,]) %>% unlist
  map28 <- map_df(data_set[2:4], as.data.frame)
  colnames(map28) <- col_names
map28 <- map28 %>%
    slice(2:3151) %>%
    unnest()
  title1 <- unlist(data_set[[1]][2,2])
  title1 %>% unlist
map28 <- map28 %>% 
  mutate(map = title1) %>%
  select(Period, `CCG code`, `CCG name`, Rate:map)


## Map 29 - this differs - only 2 subtables
tables <- split_sheet(workbook$sheets$`29`)
glimpse(tables)
  data_set <- map(tables, function(x) x$values())
  col_names <- (data_set[[2]][1,]) %>% unlist
  map29 <- map_df(data_set[2:4], as.data.frame)
  colnames(map29) <- col_names
map29 <- map29 %>%
    slice(2:3151) %>%
    unnest()
  title1 <- unlist(data_set[[1]][2,2])
  title1 %>% unlist
map29 <- map29 %>% 
  mutate(map = title1) %>%
  select(Period, `CCG code`, `CCG name`, Rate:map)

## Map 30 - this differs - only 2 subtables
tables <- split_sheet(workbook$sheets$`30`)
glimpse(tables)
  data_set <- map(tables, function(x) x$values())
  col_names <- (data_set[[2]][1,]) %>% unlist
  map30 <- map_df(data_set[2], as.data.frame)
  colnames(map30) <- col_names
map30 <- map30 %>%
    slice(2:421) %>%
    unnest()
  title1 <- unlist(data_set[[1]][2,2])
  title1 %>% unlist
map30 <- map30 %>% 
  mutate(map = title1) %>%
  select(Period = Year, `CCG code`, `CCG name`, Rate = Percent, `95% lower`:map)

## Map 36 - this differs - only 2 subtables
tables <- split_sheet(workbook$sheets$`36`)
glimpse(tables)
  data_set <- map(tables, function(x) x$values())
  col_names <- (data_set[[2]][1,]) %>% unlist
  map36 <- map_df(data_set[2], as.data.frame)
  colnames(map36) <- col_names
map36 <- map36 %>%
    slice(2:421) %>%
    unnest()
  title1 <- unlist(data_set[[1]][2,2])
  title1 %>% unlist
map36 <- map36 %>% 
  mutate(map = title1) %>%
  select(Period = Year, `CCG code`, `CCG name`, Rate = Percent, `95% lower`:map)

## Put it all together - bit of surgery on map 16 first

map16 <- map16 %>%
  mutate(Period = as.character(Period)) %>%
  select(Period:`CCG name`, Rate = ISR, `95% lower`:map)

map_all <-bind_rows(map1, map2, map3, map16, map17, map19, map20, map22, map23, map24, map25, map26, map27, map28, map29, map30, map36)

head(map_all)

```

After much effort we now have a data frame of the CCG data. There are several different sets of time periods. These will need to be recoded. We have quarterly, annual and three yearly data.

```{r, cache=TRUE}
map_all <- map_all %>%
  mutate(Period = recode(Period, 
                         "506" = "2005/06",
                         "607" = "2006/07", 
                         "708" = "2007/08", 
                         "809" = "2008/09", 
                         "910" = "2009/10", 
                         "1011" = "2010/11",
                         "1112" = "2011/12", 
                         "1213" = "2012/13", 
                         "1314" = "2013/14", 
                         "1415" = "2014/15", 
                         "2013/14_2013/14" = "2013/14", 
                         "2014/15_2014/15" = "2014/15", 
                         "2015/16_2015/16" = "2015/16"))

summary(map_all)

map_all %>%
  DT::datatable()
```


