---
title: "Changepoint analysis"
output:
  html_document: 
    toc: yes
  html_notebook: default
bibliography: changepoint.bib  
  
---
```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE)
```

## What is changepoint analysis

Changepoint analysis is...

More details here...<https://en.wikipedia.org/wiki/Change_detection>, <http://www.variation.com/cpa/tech/changepoint.html>

In R with `changepoint` package[@Killick2013]

```{r, message=FALSE, warning=FALSE}
if(!require("changepoint"))install.packages("changepoint")
   library(changepoint)

## using the example given

set.seed(1) ## random start point

x <- c(rnorm(100, 0, 1), rnorm(100, 0, 10)) ## some random data

test <- cpt.var(x) ##look for changes in variance

plot(test) ## plot

cp1 <- test@cpts ## locate changepoints

cp1

```

## Real data 1. Under 18 conceptions

```{r, message=FALSE, warning=FALSE}
library(fingertipsR)
library(dplyr)
if(!require("ggfortify"))install.packages("ggfortify")
library(ggfortify)

inds <- indicators(ProfileID = 19) ## get details of PHOF indicators

teen <- inds %>% filter(stringr::str_detect(IndicatorName, "conceptions")) ## identify the ID for conceptions data

data <- fingertips_data(IndicatorID = 20401 )  ## download data

data %>%
  filter(AreaName == "England", Category == "") %>%
  select(Timeperiod, Value) %>%
  mutate_all(funs(as.numeric)) -> test1 ## prepare data for cp analysis

rownames(test1) <- test1$Timeperiod

cp <- cpt.mean(test1[,2], method = "PELT") ## run cp on mean values/ choose PELT method


cps <- cp@cpts ## dates of change points

test1[cps,] ## filter dataset to identify cp locations

autoplot(cp) + labs(y = "rate", x = "date") ## plot



```

## Real data 2. Monthly prescribing

```{r}


## Use Indicator 92511 - Monthly prescribing data

ind_ab <- fingertips_data(IndicatorID = 92511, AreaTypeID = 19 ) %>%
  filter(AreaName == "England") %>%
  select(Timeperiod, Value) %>%
  mutate_if(is.factor, as.character) %>%
  mutate(date = seq_along(Timeperiod))

rownames(ind_ab) <- ind_ab$Timeperiod

## Create time series starting with Jan 2011 by month

ind_ab_ts <- ts(ind_ab$Value, start = c(2011, 1), frequency = 12, names = ind_ab$Timeperiod)

qplot(data=ind_ab, x = date, Value, geom = c("line", "point"),  group = 1) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

cp_pres <- cpt.mean(ind_ab_ts)
plot(cp_pres)

cp_1 <- cp_pres@cpts ## dates of change points

ind_ab[cp_1,] ## filter dataset to identify cp locations

autoplot(cp_pres) + labs(y = "rate", x = "date") + geom_line(aes(y = ind_ab$Value)) ## plot

class(ind_ab_ts)

```

```{r}
## try Bayesian version

if(!require("bcp"))install.packages("bcp")
library(bcp)

bcp1 <- bcp(ind_ab$Value, return.mcmc = TRUE)

residuals(bcp1)

plot(bcp1)

```

## Daily data: Fingertips hits

```{r}
library(RGA)

ga_token <- authorize(client.id = "934830359575-tbflfk33ce99kgatrte3ecpt4gj5kdhf.apps.googleusercontent.com", client.secret = "l5BP84XI1I5d0qrC3f21s9Gp", cache = TRUE)

## extract profile ids, website urls and site names

  lookup <- list_profiles() %>%
    select(name,  id, websiteUrl)
  
## select Fingertips id
  
  lookup %>%
    filter(stringr::str_detect(name, "Finger")) %>%
    select(id)
  
  ## download page view data
  
   df <- data.frame()

for(i in 1:length(id)) {
  id<-"78054916"
  first <- firstdate(id)
  ga <- get_ga(id, start.date = first, end.date = "today",
            metrics = "ga:users,
            ga:sessions,
            ga:avgTimeonPage,
            ga:pageviews",
            dimension = "ga:date" )
  ga <- cbind(id, ga)
  df <- rbind(df, ga)
}


## Lets just plot

qplot(data = df, date, pageviews, geom = c("line", "smooth")) + 
  labs(title = "Fingertips daily page views")
   
```

#### Changepoint analysis

```{r}
df_ts <- ts(df$users, name = df$date, start = c(2013, 330), frequency = 365.25)

cp_df_mean <- cpt.mean(df_ts, method = "BinSeg", Q = 10)
cp_df_var <- cpt.var(df_ts, method = "PELT", Q = 5)

autoplot(cp_df_mean)
 
cpdf <- cp_df_mean@cpts

df[cpdf, ] %>%
  ggplot(aes(date, pageviews)) +
  geom_line() +
  geom_point() +
  labs(title = "Changepoints for Fingertips")

```

The analysis shows that:

* Page views significantly increased every year from 
  + an average 450 a day from Oct 2014 to March 2016
  + 2600 on average since May 2016



## References

