---
title: "Rapid analysis of email feedback for Fingertips via profilefeedback mailbox"
output:
  html_notebook:
    number_sections: yes
    toc: yes
    toc_float: yes
  html_document:
    number_sections: yes
    toc: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, cache = TRUE, warning = FALSE, message = FALSE)

```



```{r}

library(readr)
library(dplyr)
library(lubridate)

```

#Introduction

Fingertips feedback is received via a number of channels including email to the [profilefeedback mailbox]("mailto:profilefeedback@phe.gov.uk).

Since 2014 there have been ~ 3700 emails to and from the mailbox. The mailbox is administered by the Fingertips team.

This short report presents a brief analysis of the emails received by the feedback. A more detailed report is available on request.

# Method

A database of emails for analysis was created as follows


* Select fields in Outlook pane
* For this analysis just using from, to , date and subject fields
* *For full email addresses and formatted body text need to *export* relevant fields*
* Copy and paste to Notepad
* Import into R as tab-separated file

```{r}
## Import
emails <- read_tsv("ftips_email.txt")

emails <- emails %>%
    mutate(date = lubridate::dmy(Received))
```

# Results

## How many emails
There were `r dim(emails)[1]` received overall between `r min(emails$date, na.rm = TRUE)` and `r max(emails$date, na.rm = TRUE)`.

## Quick look
```{r}
emails %>%
    sample_n(10) %>%
    knitr::kable()
```


## Count daily emails

The daily traffic through the mail box has increased over time.


```{r}
emails %>%
    group_by(date) %>%
    count() %>%
    arrange(date)

### and plot daily traffic

library(ggplot2)

emails %>%
    group_by(date) %>%
    count() %>%
    arrange(date)%>%
    ggplot(aes(date, n)) +
    geom_line() +
    geom_point() +
    geom_smooth(method = "gam") +
    labs(y = "Email count", 
         title = "Daily email traffic to profilefeedback mailbox")

```

## Predicted traffic

```{r, message=FALSE, warning=FALSE}

library(prophet)

email_proj <- emails %>%
    group_by(date) %>%
    count() %>%
    arrange(date) %>%
    rename(ds= date, y = n) %>%
    na.omit()

m <- prophet(email_proj)

future <- make_future_dataframe(m, periods = 365)
head(future)


summary(m)
m$growth

forecast <- predict(m, future)
m$changepoints

prophet_plot_components(m ,forecast)


```



```{r day}

library(lubridate)
library(ggmosaic)


emails <- emails %>%
    mutate(year = year(date), 
           month = month(date, label = TRUE),
           day = wday(date, label = TRUE)) 

# emails %>%
#     filter(!is.na(year)) %>%
#     group_by(year, day) %>%
#     count() %>%
#     ggplot(aes(year)) +
#     geom_bar(aes(fill = day, stat = n)) 

```

```{r}
emails2015 <- emails %>%
    filter(date >= "2015-01-01" )
```


## What topics?

We can plot a word cloud to show the relative frequency of feedback for different profiles.

```{r, fig.height=6, fig.width=6}
library(stringr)
library(tidytext)


library(wordcloud)

words_body <- emails2015 %>%
    select(Subject) %>%
    unnest_tokens(word, Subject) %>%
    anti_join(stop_words)%>%
    count(word, sort = TRUE) %>%
    filter(!word %in% c("profile", "profiles", "fw", "feedback"))




with(words_body, wordcloud(word, n, min.freq = 5, max.words ='INF', 
                           scale = c(8, 0.1),rot.per = 0.5, random.order = FALSE, colors = brewer.pal(8, "Dark2")))
```


## API mentions

We launched the Fingertips automated programming interface (API) in November 2016, which is beginning to receive attention and feedback.


```{r}
emails %>%
    filter(str_detect(Subject, "[Aa][Pp][Ii]")) %>%
    group_by(year, month) %>%
    count()
```
```{r}
users <- emails2015 %>% select(From) %>% distinct %>% nrow %>% list


```


### Who sends?

In total `r users` people have used the mail box. The users with more than one email are:

```{r}

emails %>%
    group_by(From, year) %>%
    count() %>% 
    arrange(-n) %>%
    filter(n>1)%>%
    tidyr::spread(year, n, fill = 0) %>%
    knitr::kable()
```

