---
title: 'Fingertips user: Google analytics report'
author: "Julian Flowers"
date: '`r Sys.Date()`'
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(govstyle)
library(data.table)
options(digits = 1)
##source("updated_ga.R")
```

```{r data download and convert to data.table}
ds <- fread("ga_fingertips.csv")


dsdt <- as.data.table(ds)

## Index
setkey(dsdt, date)

## Calculated fields

dsdt <- dsdt[, hits_per_session := hits/sessions]
dsdt <- dsdt[, date := lubridate::ymd(date)]

rm(ds)
```


```{r}

mindate <- range(dsdt$date)[1]
maxdate <- range(dsdt$date)[2]

```

# The report

This report covers Google analytics for Fingertips from `r mindate` to `r maxdate`.



## Trend in hits

```{r}

dsdt[, .(total_hits = sum(hits)), by = date] %>%
  ggplot(aes(date, total_hits)) +
  geom_line(aes(group = 1)) +
  geom_smooth(aes(group = 1)) +
  theme_gov() +
  theme(axis.text.x = element_text( size = rel(0.6) ))+ 
  labs(title = "Trend in Fingertips: total hits") +
  scale_x_date(name = "Date", date_breaks = "3 month", date_labels = "%m/%Y")


```


## Trend in sessions

```{r}

dsdt[, .(total_sessions = sum(sessions)), by = date] %>%
  ggplot(aes(date, total_sessions)) +
  geom_line(aes(group = 1)) +
  geom_smooth(aes(group = 1)) +
  theme_gov() +
  theme(axis.text.x = element_text( size = rel(0.6) ))+ 
  labs(title = "Trend in Fingertips: daily sessions") +
  scale_x_date(name = "Date", date_breaks = "3 month", date_labels = "%m/%Y")




```

## Trend in hits per session

```{r}

dsdt[, .(hits_sessions = sum(hits)/ sum(sessions)), by = date] %>%
  ggplot(aes(date, hits_sessions)) +
  geom_line(aes(group = 1)) +
  geom_smooth(aes(group = 1)) +
  theme_gov() +
  theme(axis.text.x = element_text( size = rel(0.6) ))+ 
  labs(title = "Trend in Fingertips: daily hits per session ratio", 
       y = "Hits per session") +
  scale_x_date(name = "Date", date_breaks = "3 month", date_labels = "%m/%Y")





```

## Trend in event categories

```{r, fig.height=6, fig.width=8, message=FALSE, warning=FALSE}

dsdt[, .(events  = sum(totalEvents)), by = c("date", "eventCategory")][events > 20] %>% 
  
  ggplot(aes(date, events)) +
  geom_line(aes(group = 1)) +
  geom_smooth(aes(group = 1)) +
  theme_gov() +
    scale_x_date(name = "Date", date_breaks = "6 months", date_labels = "%m/%Y") +

    facet_wrap(~ eventCategory, scales = "free", ncol = 4, labeller =  labeller(label_wrap_gen(width = 20))) +
  theme(strip.text.x = element_text(size  =rel(.6)) ) +
  labs(title = "Trend in Fingertips: events")
```


```{r}
options(scipen = 2)
days <- as.numeric(maxdate - mindate)
prac <- dsdt[eventCategory == "PracticeSearch",.(mean = sum(totalEvents)/days) ] 
search <- dsdt[eventCategory == "Search",.(mean = sum(totalEvents)/days) ] 
download <- dsdt[eventCategory == "Download",.(mean = sum(totalEvents)/days) ] 


```

On average there are:

* `r prac$mean` searches for GP data per day
* `r search$mean` searches per day
* `r download$mean` downloads per day


## Trend in downloads

```{r}

library(ggfortify)
library(changepoint)
downloads <-dsdt[eventCategory == "Download", .(tot_downloads = sum(totalEvents)), by = date ] 

downloads.ts <- as.ts(downloads$tot_downloads, start(2016, 1), frequency = 365.25 )

autoplot(cpt.mean(downloads.ts)) +
  labs(title = "Changepoint anlaysis of downloads")
  
cpt <- cpt.mean(downloads.ts)
cp <- downloads[cpt@cpts, ][1]


precp <- downloads[date< "2017-05-01", .(mean_downloads = sum(tot_downloads)/ .N)]
postcp <- downloads[date >= "2017-05-01", .(mean_downloads = sum(tot_downloads)/ .N)]
```

  
* Daily downloads significantly increased after implementation of download options `r cp$date` from a daily average of `r precp$mean_downloads` downloads to `r postcp$mean_downloads`.


## Use by city - top 50

```{r, fig.height=6, fig.width=8, message=FALSE, warning=FALSE, results="asis"}
n <- 50

dsdt[, .(total_hits = sum(hits)), by = city][order(-total_hits)] %>%
  top_n(n) %>%
  ggplot(aes(reorder(city, total_hits), log10(total_hits))) +
  geom_col(fill = gov_cols[10]) +
  coord_flip() +
  theme_gov() +
  geom_text(aes(label = total_hits, hjust = 0), size = rel(.8)) +
  labs(title = paste("Top", n, " cities") , 
        y = paste("Total sessions since", mindate, "(log scale)"), 
        x = "")


```

## Top 50 landing pages


```{r, fig.height=6, fig.width=8, message=FALSE, warning=FALSE}



dsdt[, .(total_sessions = sum(sessions)), by = landingPagePath][order(-total_sessions)] %>%
  top_n(n) %>%
  mutate(page = stringr::str_replace(landingPagePath, "fingertips.phe.org.uk/profile/", ""), 
         page = stringr::str_replace(page, "/", "")) %>%
  ggplot(aes(reorder(page, total_sessions), log10(total_sessions))) +
  geom_col(fill = gov_cols[10]) +
  coord_flip() +
  geom_text(aes(label = total_sessions), size = rel(1.5), hjust = 0) +
  theme_gov() +
  labs(title = paste("Top", n, "landing pages"), 
       y = paste("Total sessions since", mindate, "(log scale)"), 
       x = "")


```

## Top 50 searches

```{r}
dsdt[stringr::str_detect(landingPagePath, "search"), .(total_sessions = sum(sessions)), by = landingPagePath][order(-total_sessions)] %>%
  mutate(page = stringr::str_replace(landingPagePath, "fingertips.phe.org.uk/search/", ""), 
         page = stringr::str_replace(page, "/", "")) %>%
  filter(!stringr::str_detect(page, "^finger")) %>%
  select(page, total_sessions)%>%
  top_n(50) %>%
ggplot(aes(reorder(page, total_sessions), total_sessions)) +
  geom_col(fill = gov_cols[10]) +
  coord_flip() +
  geom_text(aes(label = total_sessions), size = rel(2), hjust = 0 ) +
  theme_gov() +
  theme(axis.text = element_text(size = rel(0.5))) +
  labs(title = paste("Top", n, "Fingertips searches"), 
       y = paste("Total searches since", mindate), 
        x = "")
```

## Searches by city

```{r}
dsdt[stringr::str_detect(landingPagePath, "search"), .(total_sessions = sum(sessions)), by = .(landingPagePath, city)][order(-total_sessions)] %>%
  mutate(page = stringr::str_replace(landingPagePath, "fingertips.phe.org.uk/search/", ""), 
         page = stringr::str_replace(page, "/", "")) %>%
  filter(!stringr::str_detect(page, "^finger")) %>%
  select(page, total_sessions)%>%
  top_n(100)
```

