---
title: "GP prevalence estimates from Imperial College"
author: "Julian Flowers"
date: "14 March 2017"
output:
  html_document:
    toc: yes
  word_document:
    fig_caption: yes
    toc: yes
---
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
 
Sys.setenv(http_proxy="http://158.119.150.18:8080")
Sys.setenv(https_proxy="https://158.119.150.18:8080")
```
 
# Introduction
 
This short report presents a brief analysis of the GP level prevalence estimates fom Imperial College provided to PHE. There are estimates for:
 
* CHD (age 55-79)
* stroke (age 55 - 79)
* peripheral arterial disease (age 55 -79)
* depression
* COPD
* hypertension - diagnosed and undiagnosed
 
The method for calculating these estimates are described elsewhere.
 
The data is available at...
 
```{r, include=FALSE, warning = FALSE, message= FALSE}
library(readr)
library(ggplot2)
library(tidyr)
library(dplyr)
library(ggmap)
##library(container)
 
 
 
setwd("~/Prev_est")
data <- read_csv("2017-03-14 gp_prev_ests.csv")
data_qa <- read_csv("gp.estimates2017-03-14.csv")
options(digits = 3)
 
```
 
The maximum number of practices for which estimates are available is `r nrow(data_qa)`, but for some diseases there are fewer. The counts are shown in the table.
 
 
 
```{r}
## count practices in `data_qa`
 
data_qa %>%
    purrr::map_df(function(x) sum(!is.na(x))) %>%
    select(practice_code, contains("estimate")) %>%
    gather(disease, count, 2:9) %>%
    select(-practice_code) %>%
    knitr::kable(format = "pandoc",
                 caption = "Practice counts by disease")
   
 
```
 
 
 
## Summary statistics
 
Hypertension and depression have the highest prevalence in this group of disase estimates.
 
```{r}
data %>%
    group_by(disease.x) %>%
    summarise(practices_count = n(),
              meanprev = mean(prevalence, na.rm = TRUE),
              sdprev = sd(prevalence, na.rm = TRUE),
              medianprev = median(prevalence, na.rm = TRUE),
              minprev = min(prevalence, na.rm = TRUE),
              maxprev = max(prevalence, na.rm = TRUE),     
              range = max(prevalence, na.rm = TRUE) - min(prevalence, na.rm = TRUE),
              IQR = quantile(prevalence, probs = 0.75, na.rm = TRUE) - quantile(prevalence, probs = 0.25, na.rm = TRUE))
 
                
 
```
 
We can show the distribution of prevalences for each disease as distrbution plots.
 
```{r}
data %>%
    ggplot(aes(disease.x, prevalence, fill = disease.x)) +
    geom_violin(draw_quantiles = c(0.1, 0.5, 0.9)) +
    labs(y = "Prevalence (%)",
         x = "",
         title = "Violin plots of variation in practice level disease prevalence") +
    scale_y_log10()
   
    
    
```
 
And plot as maps
 
```{r}
gpdata <- read_delim("C:/Users/julian.flowers/Downloads/GP.csv",  "\t", escape_double = FALSE, trim_ws = TRUE)
 
gpdata <- gpdata %>% clean_names()
 
gddata1 <- gpdata %>%
    left_join(data_qa, by = c("organisationcode" = "practice_code"))
   
 
loc <- gpdata %>% select(latitude, longitude)
loc <- slice(loc, 5:30)
##mymap <- get_map("London")
##ggmap(mymap) + geom_jitter(aes( y = latitude, x= longitude), colour = organisationtype,
                          data = gpdata, size = 1, alpha = 0.5
                          ) + coord
```
 
 