---
title: "Peer benchmark tool UTLA data"
author: "Julian Flowers"
date: "29 July 2016"
output:
  html_notebook: 
    number_sections: yes
    toc: yes
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE, cache=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

#Analysis of the PHE peer benchmark tool - profiling the profile

The Peer Benchmarking Tool contains a selection of indicators that reflect PHE’s national priorities. They provide a summary of determinants, risk factors and outcomes using indicators available in other Fingertips tools - particularly the Public Health Outcomes Framework. They are broken down into the following categories:

* Giving children and young people the best start in life (CYP)
* Mental health and wellbeing (MH&WB)
* Drugs and alcohol (D&A)
* Sexual health (SH)
* Obesity
* Smoking
* Dementia
* Health inequalities (HI)
* Health checks (HC)
* PHE's core offer to the NHS (CO)

The excel file can be downloaded from [Fingertips](http://fingertips.phe.org.uk/hub-tool), and the indicator list [here](http://fingertips.phe.org.uk/profile/hub-tool/data#gid/1938132836/ati/102).


```{r Load libraries, message=FALSE, warning=FALSE, include=FALSE}

suppressPackageStartupMessages(c(
library(viridis),
library(ggplot2),
library(knitr),
library(readxl),
library(corrplot),
library(tidyr),
library(dplyr),
##library(networkD3)
library(data.table),
library(DT),
library(cluster),
library(NbClust),
library(openxlsx)
))
```

##Read in data

Reading in data can done via the `readxl` package. UA data are in sheet 3. 

```{r Import data and change headers to lower case, message=FALSE, warning=FALSE, cache=TRUE}
setwd("~/Documents/R_projects/profiles")

profile <-read_excel("peerdata.xlsx", 3)

profile <- as.data.table(profile) ## convert to data.table

colnames(profile) <- tolower(make.names(colnames(profile))) ## tidy up column names
```

##Review data 

We will extract key variables and create an interactive data table to help explore the data and enable the data to easilty downloaded.

```{r Datatable, message=FALSE, warning=FALSE}
profileDT <- profile[, .(indicator, time.period, area.name, sex, value = round(value,2))] ## extract relevant columns and round data

u <- unique(profileDT$indicator) ## view indicator names

## Reclassify some variables into existing groups

u[106:117] <- paste("Smoking", unique(profileDT$indicator)[106:117], sep = ": ")

u[118:131] <- paste("Alcohol", unique(profileDT$indicator)[118:131], sep = ": ")

u[87:98] <- paste("Other", unique(profileDT$indicator)[87:98], sep = ": ")

l <- levels(factor(profileDT$indicator))

recoder::recoder(l,u)

hubprof <- tidyr::separate(profileDT, indicator, c("category", "indicator"), sep = ":") ## split indicator name from category

hubprof ## top 1000 rows

unique(hubprof$category)





# datatable(profileDT, class = "compact", caption = "Hub dataset",extensions = 'Buttons', options = list(
#     dom = 'Bfrtip',
#     buttons = c('copy', 'csv', 'excel', 'pdf', 'print'))) ## create datatable to explore the data values
```

##Extract latest data only and shorten indicator names

First create a unique indicator index by combining indicator names, time periods and gender breakdown

```{r Extract latest data, message=FALSE, warning=FALSE}
profile <- profile %>% mutate(indshort = substring(indicator, 1, 35),indextime = paste(indicator, age, sex,  time.period, sep = "-"), indgend = paste(indshort, sex, sep = '-')) %>%
                             arrange (desc(time.period))
profile <- as.data.table(profile)
proflatest <- profile[, .SD[1], by = indgend]

profile1 <- profile[indextime %in% proflatest$indextime,.(indextime, indicator, indshort, indgend, time.period, area.name, value, lower.ci, upper.ci, count, denominator)]

dim(profile1)

```


## Exploration

```{r Explore, fig.height=13, fig.width=7, message=FALSE, warning=FALSE}


length(unique(profile1$indshort)) ## 111 indicators

length(unique(profile1$indextime)) ## the dataset contains 160 indicator - gender combinations


count(profile1, time.period) %>% arrange(desc(n)) ## there are 14 different 'latest' time periods

count(profile1, indgend, time.period) %>% arrange(desc(n)) 

profile1 <- profile1[!area.name %in% c("City of London", "Isles of Scilly"),]

count(profile1, indgend, time.period) %>% arrange(desc(n))

## boxplots

qplot(data = profile1, indgend, log(value), geom = "boxplot", fill = time.period) + coord_flip() + theme(axis.text.y = element_text(size = 6)) + xlab("") + theme(legend.position = "bottom")

```


# Missing data

Data in these profiles can be missing because of:

* Poor data quality
* Data not collected
* Suppression due to small numbers

It tends not to be missing at random (MAR). 

Some modelling approaches such as k-means analysis don't work well with missing data so decisions need to be made about tackling it.

```{r Missing data}

## 

## How much is missing?
mean(is.na(profile1)) ## ans ~ 4%

## Where is it missing?
## By indicator
na <- profile1 %>% group_by(indgend) %>% summarise(na = 100 * mean(is.na(value))) %>% filter(na >20) %>% arrange(desc(na))

kable(na) ## 5 indicators have > 20% data missing - they will be excluded from further analysis. 35 have missing data  <20% - will impute missing values

## By area

na1 <- profile1 %>% group_by(area.name) %>% summarise(na.a = 100 *
mean(is.na(value))) %>% filter(na.a >0) %>% arrange(desc(na.a))
 ## all areas have at least some missing data. It is almost 20% in Rutland.

## There are lots of approaches to imputing missing data. For simplicity will replace missing data with the mean values for each indicator

## Exclude indicators with >20% missing data

highmiss <- c("Dementia: Estimated diagnosis rate -Persons",
              "MH&WB: Suicide rate-Female", 
              "CO: Tuberculosis: Treatment complet-Persons",
              "MH&WB: Self-reported well-being - p-Persons",
              "CO: Preventable sight loss - diabet-Persons")
profile1 <- profile1[!indgend %in% highmiss, ] 

## Impute


profile1.imp <- profile1 %>% group_by(indgend) %>% mutate(newval = ifelse(is.na(value), mean(value, na.rm = TRUE), value))

mean(is.na(profile1.imp$newval)) ## check it is 0

## Calculate mean values for each indicator

meanprof <- profile1 %>% group_by(indgend) %>% summarise(meanval = mean(value, na.rm = TRUE))

meanprof1 <- profile1.imp %>% group_by(indgend) %>% summarise(meanval1 = mean(newval, na.rm = TRUE))

qplot(meanprof$meanval, meanprof1$meanval1) + geom_smooth(method = "lm")

## shows that mean values with and without missing data are virtually identical

```


# Correlations

```{r Correlation, fig.height=12, fig.width=12, message=FALSE, warning=FALSE}

## Select fields

profw <- profile1.imp %>% select(indgend, area.name, value)

## Reshape dataset

require(tidyr)

profw <- profw %>% spread(indgend, value)
dim(profw)
mean(is.na(profw))

## remove nas
profw1 <- na.omit(profw)
require(corrplot)
cor<- cor(profw1[, -1])
corrplot(cor, tl.cex = .5, addgrid.col = NA, method = "square", tl.col= 'black', order = "hclust", addrect = 10, title = "Correlations between all indicators in the Peer benchmarking tool")

## ggcorrplot version

require(ggcorrplot)


ggcorrplot(cor,  method = "square", colors = c("red", "white", "blue"), hc.order = TRUE, outline.color = "white", type = "upper", hc.method = "complete", tl.cex = 5, tl.col = "darkgrey") + ggtitle("Correlations between all indicators in the \nPeer benchmarking tool")

```

There is a large group of indicators which are highly correlated

```{r D3 heatmap, fig.height=9, fig.width=7, message=FALSE, warning=FALSE}
library(d3heatmap)
profw <- data.frame(profw)


rownames(profw) <- profw$area.name
d3heatmap(profw[, -1], xaxis_font_size = "5pt", yaxis_font_size = "5pt", k_row = 7, k_col = 8)

```

## k means

Looks like  ~7 clusters would work...NB Lewisham
Need to scale the dataset

```{r, fig.height=10, fig.width=10}
## Need to remove missing data

profw.k <- apply(profw[, -1], 2, function(x) ifelse(is.na(x), mean(x, na.rm = TRUE), x))

profw.k <- as.data.frame(unlist(profw.k))
rownames(profw.k) <- profw$area.name

str(profw.k)

k <- kmeans(scale(profw.k), 7, nstart = 25)

k$size

k$centers

k$cluster

profw.k$cluster <- k$cluster

c <- data.frame(cbind(scale(profw.k), k$cluster))
rownames(c) <- profw$area.name


## Cluster lookup table

c1 <- data.frame(profw$area.name, c[, 157])
names(c1) <- c("Local Authority","Cluster")
require(DT)
datatable(c1, options = list(pageLength = 25))

## reshape
p <- gather(c[, -156], ind, value, 1:155)
dim(p)
str(p)

## plot indicators...

kplot1 <- ggplot(p, aes(substring(ind,1, 12), value, colour = factor(V157))) + geom_line(aes(group = factor(V157)))
kplot1 + facet_wrap(~V157) + coord_polar() + geom_hline(yintercept = 0) + ggtitle("Cluster profiles")

```

