---
title: "K&I User statistics"
output:
  pdf_document: 
    latex_engine: 
  html_notebook: default
  html_document: default
  word_document: default
  
  
---

`r paste("Date run:", lubridate::today())`

This report presents summary statistics for 10 K&I websites which currently account for the majority of web hits for K&I products.
The sites are:

- [Fingertips](http://fingertips.phe.org.uk)
- [PHOF](http://www.phoutcomes.info)
- [NOO](http://www.noo.org.uk)
- [CHimat](http://www.chimat.org.uk)
- [Local health](http://www.localhealth.org.uk)
- [YHPHO](http://www.yhpho.org.uk)
- [Healthier lives](http://healthierlives.phe.org.uk)
- [End of life intelligence network](http://www.endoflifecare-intelligence.org.uk/home)
- [Tobacco profiles](http://www.tobaccoprofiles.info)
- [SHAPE](https://shape.phe.org.uk/)

(Note: PHOF, Healthier lives and Tobacco profiles are built on the Fingertips platform)

The report shows:

*Summary statistics*

1. Hits for yesterday, 30 days ago, 365 days ago, and % change from this day last year for each site
2. Hits for the last month and the last year for each site 
3. Global hits for all sites combined


*Heat maps and charts* 

1. Weekly Google searches. 
2. Searches and downloads on Fingertips and PHOF
3. Daily hits since April 2013 represented as a 'calendar' heatmap
4. Dwell time

*Highlights*

* Collective use of the 'top ten' K&I sites is about 25,000 hits and 10 person days per day (ie equivalent of 10 people spending 24 hours on the sites each day - or 1000 people spending 15 minutes)
* There has been considerable increase in total time on sites over the last year
* The most popular sites are NOO and Fingertips
* Hits and dwell time on Fingertips has grown rapidly in the last year as the number of profiles avaiable has grown - now 64.



```{r message=FALSE, warning=FALSE, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)

suppressPackageStartupMessages(c(library(data.table),
  library(dplyr) ,        # data manipulation
  library(lubridate),    # date manipulation
  library(tidyr),      # reshaping
  library(RGA),          # access to Google analystics API
  library(gridExtra),    # chart layout
  library(ggplot2),      # charting
  library(viridis),
  library(ggfortify),     # charting extras
  library(ggTimeSeries)))  # charting extras
```


```{r}
## authorise access to google API
ga_token <- authorize(client.id = "934830359575-tbflfk33ce99kgatrte3ecpt4gj5kdhf.apps.googleusercontent.com", client.secret = "l5BP84XI1I5d0qrC3f21s9Gp", cache = TRUE)

## extract profile ids, website urls and site names

  lookup <- list_profiles() %>%
    select(name,  id, websiteUrl)
  
```

```{r extract users and pageviews, cache= TRUE, include=FALSE}
library(lubridate)
ids <- c("12177926","32366919","35885149","17588911","102894568","35140940","47898340","65073818","68505331","78054916","93017633")

lookup1 <- lookup %>% filter(id %in% ids)


  
  ## extract data and create data table
  ##ids <- as.character(lookup1[,"id"])
  df <- data.frame()

for(i in 1:length(ids)) {
  id<-ids[i]
  first <- firstdate(id)
  ga <- get_ga(id, start.date = first, end.date = "today",
            metrics = "ga:users,
            ga:sessions,
            ga:avgTimeonPage,
            ga:pageviews",
            dimension = "ga:date" )
  ga <- cbind(id, ga)
  df <- rbind(df, ga)
}
  
df <- df %>% 
  left_join(lookup1) %>% 
  unite("webname", name:websiteUrl, sep = "_")

```

```{r, cache=TRUE, include=FALSE}
ids <- c("65073818","78054916") 
  ## extract data and create data table
  ##ids <- as.character(lookup1[,"id"])
 events <- data.frame()

for(i in 1:length(ids)) {
  id<-ids[i]
  first <- firstdate(id)
  ga <- get_ga(id, start.date = first, end.date = "today",
            metrics = "ga:totalEvents",
            dimension = "ga:date, ga:eventCategory", fetch.by = "month" )
  ga <- data.frame(cbind(id, ga))
  events <- bind_rows(events, ga)
}

## convert to data_frame format

events <- events %>%
  left_join(lookup1)

```
```{r, cache=TRUE}
## Extract specific date values

options(digits = 2)

df1 <- df %>% 
  filter(!webname == "www.improvinghealthandlives.org.uk_http://www.improvinghealthandlives.org.uk") %>%
  mutate(yesterday = ifelse(date == lubridate::today() - 1, 1, 0), 
         last_month = ifelse(date == lubridate::today() - 30, 1, 0), 
         last_year = ifelse(date == lubridate::today() - 365 , 1, 0), 
         lastmonth = ifelse(date > lubridate::today() - 30, 1, 0), 
         lastyear = ifelse(date >= lubridate::today() - 365, 1, 0)) %>% 
  select(webname, date, pageviews,  yesterday:lastyear) %>%
  gather(period, views, yesterday:lastyear) %>%
  filter(views == 1) %>%
  group_by(webname, period) %>%
  summarise(views = sum(pageviews)) %>%
  spread(period, views) %>%
  select(webname, Yesterday = yesterday, `30 days ago` = last_month, 
         `A year ago` = last_year, 
         `Last month` = lastmonth, 
         `Total last 365 days` = lastyear) %>%
  mutate(`Year on year change (%)` = 100 * (Yesterday - `A year ago`)/ `A year ago`) %>%
  arrange(-`Total last 365 days`) %>%
  as.data.frame()

  rownames(df1) <- c("Fingertips", "NOO", "CHimat",  "Local health", "PHOF", "YHPHO", "End of Life", "Healthier Lives", "Tobacco", "SHAPE")
  
  df1 <- df1 %>% select(-webname) 
  

```

```{r, echo=FALSE}
knitr::kable(df1, format = "markdown", digits = 2)

```

```{r Searches, fig.width = 8, fig.align= "center"}

events %>%
  filter(eventCategory %in% c("Search", "Download")) %>%
  ggplot(aes(date, totalEvents)) +
  geom_line(aes(colour = name)) +
  geom_smooth(aes(group = name)) +
  facet_wrap(~eventCategory) +
  labs(y = "Daily events", 
       x = "Date", 
       title = "Daily searches and downloads from Fingertips and PHOF sites")
 
```

```{r fig.height=6, fig.width=6, message=FALSE, warning=FALSE}
options(scipen = 2)

df %>%
  mutate(total_page_time = avgTimeonPage*pageviews) %>%
  group_by(date) %>%
  summarise(`Dwell time` = sum(total_page_time)) %>%
  ggplot(aes(date, `Dwell time`)) +
  geom_line(colour = 'lightgrey', alpha = 0.3) + 
  geom_smooth() +
  labs(title = "Combined dwell time on sites", 
       subtitle = "Minutes of use per day") +
  theme_bw()




```

```{r Total hits, fig.width = 6,fig.align="center"}
df_filt <- df %>%
  filter(date >= "2013-04-01")

df_filt %>% 
  mutate(date = lubridate::ymd(substr(date, 1, 10))) %>%
  group_by(date) %>%
  summarise(Pageviews = sum(pageviews)) %>%
  ggplot_calendar_heatmap("date", "Pageviews") +
  scale_fill_continuous(low = "red", high = "green") +
  facet_wrap(~Year, ncol = 1) +
  ggtitle("Total pageviews of K&I web products")



```

```{r PHOF, fig.width=6, fig.align="center"}

df_filt %>%
  filter(webname == "PHOF_http://www.phoutcomes.info") %>%
    mutate(date = lubridate::ymd(substr(date, 1, 10))) %>%
  group_by(date) %>%
  summarise(Pageviews = sum(pageviews)) %>%
  ggplot_calendar_heatmap("date", "Pageviews") + 
  scale_fill_continuous(low = "red", high = "green") + 
  facet_wrap(~Year, ncol = 1) +
  ggtitle("Total pageviews: PHOF")

```

```{r Fingertips, fig.width=6, fig.align="center"}


df_filt %>%
  filter(webname == "Fingertips site data_https://phe.org.uk") %>%
    mutate(date = lubridate::ymd(substr(date, 1, 10))) %>%
  group_by(date) %>%
  summarise(Pageviews = sum(pageviews)) %>%
  ggplot_calendar_heatmap("date", "Pageviews") + 
  scale_fill_continuous(low = "red", high = "green") + 
  facet_wrap(~Year, ncol = 1) +
  ggtitle("Total pageviews: Fingertips")
```

```{r Local Health, fig.width=6, fig.align="center"}

df_filt %>%
  filter(webname == "www.localhealth.org.uk_http://www.localhealth.org.uk") %>%
    mutate(date = lubridate::ymd(substr(date, 1, 10))) %>%
  group_by(date) %>%
  summarise(Pageviews = sum(pageviews)) %>%
  ggplot_calendar_heatmap("date", "Pageviews") + 
  scale_fill_continuous(low = "red", high = "green") + 
  facet_wrap(~Year, ncol = 1) +
  ggtitle("Total pageviews: Local Health")


```