---
title: "Text mining PHE's blogs"
output:
  html_notebook:
    number_sections: yes
    toc: yes
    toc_float: yes
  html_document:
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---

# Introduction

## Text is the new data

PHE produces numerous reports, bulletins, and communications, and receives large amounts of feedback. In recent years the ability to analyse text as data has developed rapidly, and there are now tools which can help us gain insight from documents and bodies of texts. These tools allow us to rapidly analyse large numbers of documents. This note applies some of these techniques to analysing Duncan Selbie's Friday Messages.

There are a number of steps:

* Automated downloading all the bulletins (web-scraping)
* Extracting and cleaning the bulletin text
* Tidying the data to make it analysable
* Applying visualisation techniques like word clouds
* Applying text mining and clustering techniques like topic modelling which can help discover the main themes of the bulletins, and automatically classify them based on these themes.

This analysis is conducted using the statistical package `R` which is rapidly becoming the main tool for undertaking this kind of work.

```{r setup, cache=TRUE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, warning = FALSE, message = FALSE)

```

First we need to load the libraries for the analysis.

```{r load libraries}
library(knitr)
suppressPackageStartupMessages(library(rvest))
suppressPackageStartupMessages(library(stringr))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(httr))
suppressPackageStartupMessages(library(tm))
suppressPackageStartupMessages(library(pdftools))
suppressPackageStartupMessages(library(tidytext))
suppressPackageStartupMessages(library(ggplot2))
library(tidyverse)
library(govstyle)

source("~/themejf.R") ## A standard theme for plots - this should be changed
```

## Downloading messages - web scraping

Then we need to get the data. This process identifies the URLs of the Public Health Matters blog. To identify the parts of the html pages the handy [selectorgadget](http://selectorgadget.com) tool is very useful.

The first step is to extract all the relevant URLs.

```{r Identify URLs, message=FALSE, warning=FALSE, cache=TRUE}


## Scraping bulletins



### This is the main stem of the URLs for Public Health Matters Blogs
url <- "https://publichealthmatters.blog.gov.uk/category/duncan-selbie-friday-message/"

page <- read_html(url)

urls <- page %>%
  
  html_nodes("a") %>%       # find all links
  
  html_attr("href")     # get the url

urls <- unique(urls[grepl("friday",urls)])  ## Select those which are friday messages
url_comment <- urls[stringr::str_detect(urls, "comments$")]
url_category <- urls[stringr::str_detect(urls, "category")]

urls <- urls[!urls %in% url_comment]
urls <- urls[!urls %in% url_category]

url1 <- "https://publichealthmatters.blog.gov.uk/category/duncan-selbie-friday-message/page/2/"

page1 <- read_html(url1)

urls1 <- page1 %>%
  
  html_nodes("a") %>%       # find all links
  
  html_attr("href")     # get the url

urls1 <- unique(urls1[grepl("friday",urls1)])  ## Select those which are friday messages
url_comment1 <- urls1[stringr::str_detect(urls1, "comments$")]
url_category1 <- urls1[stringr::str_detect(urls1, "category")]

urls1 <- urls1[!urls1 %in% url_comment1]
urls1 <- urls1[!urls1 %in% url_category1]

url2 <- "https://publichealthmatters.blog.gov.uk/category/duncan-selbie-friday-message/page/3/"

page2 <- read_html(url2)

urls2 <- page2 %>%
  
  html_nodes("a") %>%       # find all links
  
  html_attr("href")     # get the url

urls2 <- unique(urls2[grepl("friday",urls2)])  ## Select those which are friday messages
url_comment2 <- urls2[stringr::str_detect(urls2, "comments$")]
url_category2 <- urls2[stringr::str_detect(urls2, "category")]

urls2 <- urls2[!urls2 %in% url_comment2]
urls2 <- urls2[!urls2 %in% url_category2]

urls_all <- c(urls, urls1, urls2)

```

Then read the text and clean the data.


```{r Extract text and create data frame, echo=FALSE, message=FALSE, warning=FALSE}
df <- data.frame()
for(url in urls_all){

test1 <- read_html(url) %>% 
  html_nodes("p")



test2 <- html_text(test1) %>%
  str_replace_all("\n", "")

## Remove 1st row

test2 <- test2[2:length(test2)]

## Extract title

title <- test2[1]

## Extract body text

text <- test2[5:length(test2)]

text <- str_c(text, collapse = "")

## Remove repeated text at the end of each blog

test3 <- gsub("With best wishes.*", "", text )

test4 <- data.frame(test3, title)

df <- bind_rows(df, test4)

}



```


We'll extract the date from the title

```{r}

date <- stringr::str_split(df$title, pattern = "/")

year <- map(date,  4)
month <- map(date, 5)
day <- map(date, 6)

date1 <- paste(year, month, day, sep = "-")

df <- cbind(df, date1)

df$date1 <- as.Date(as.character(df$date1))

df <- rename(df, 
             text = test3)


```



### Text processing

##### Removing 'stop words'

Stop words are common English words which occur frequently in all documents and are generally removed for analytical purposes. In addition, there are words common to all bulletins which add little value in analysis - we'll add these to the `stop_words` lexicon.

```{r}


bulletin_sw <- data_frame(word = c("phe","friday", "dear", "week", "health", "published", "bulletin", "press", "public", "phe's", "www.gov.uk", "news", "publications", "gateway", "formore"), lexicon = "SMART" )

stop_words1 <- bind_rows(stop_words, bulletin_sw)
```


## Visualisation

### Words per document

We can do some simple analysis.

```{r echo=FALSE, fig.height=6, cache=TRUE}
df %>% 
  group_by(date1) %>% 
  unnest_tokens(word, text) %>%
  count(word, sort = TRUE)%>%
  summarise(words = sum(n)) %>%
  ggplot(aes(reorder(date1, words), words)) +
  geom_col(fill = "blue") +
  theme_gov() +
  coord_flip() +
  labs(x="", 
       y = "Number of words",
       title = "Number of words per blog") +
  theme(axis.text.y = element_text(size  = 10))
```


### Simple wordcloud

And wordclouds...
 
```{r echo=FALSE, fig.height=6, fig.width=6, cache=TRUE}
library(wordcloud)

cloud <- df %>% 
  unnest_tokens(word, text) %>%
  anti_join(stop_words1) %>%
  count(word, sort = TRUE)

with(cloud, wordcloud(word, n, max.words = 'INF', scale = c(8, 0.2), 
                      rot.per = 0.4, random.order = FALSE, 
                      colors = brewer.pal(8, "Dark2")))
```

```{r echo=FALSE, fig.height=8, fig.width=8}
library(wordcloud2)
library(tidyr)


cloud1 <- df %>% 
  unnest_tokens(bigram, text, token = "ngrams", n = 2) %>%
  count(bigram, sort = TRUE)


## Split the bigrams back into separate words
cloud_sep <- cloud1 %>%
  separate(bigram, c("word1", "word2"), sep = " ") 


## and filter out the 'stop words'

cloud_filt <- cloud_sep %>%
  filter(!word1 %in% stop_words1$word) %>%
  filter(!word2 %in% stop_words1$word)
  
## and recombine

cloud_new <- cloud_filt %>%
  unite(bigram, word1, word2, sep = " ")

## and replot the word cloud 


wordcloud2(cloud_new, size = 0.6)

```

This shows the main topics or themes of the messages. "Local authorities" is the most commonly mentioned topic, with "suicide prevention", "bowel cancer" and "blood pressure" amongst other common themes.

# Analysis by blog

We can extend the anlaysis further by looking at the distribution of terms in individual bulletins, and then looking for patterns to see if bulletins can be clustered according to content.

First we can look at a single bulletin:

```{r}
df%>%
  sample_n(1) %>%
  knitr::kable(format = "pandoc")
```

We can then create a per document per term table known as a Document Term Matrix (DTM). We can count the terms per document.

```{r echo=FALSE}
corp_dtm <- df %>%
  group_by(date1) %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words1) %>%
  count(date1, word, sort = TRUE) 


```

Next we can create the DTM.

```{r include=FALSE}
corp_dtm <- corp_dtm %>%
  cast_dtm(date1, word, n)


  
```


## Word associations 

We can see which words tend to appear together in the bulletins.

```{r associations}
## For example...

### Alcohol

findAssocs(corp_dtm, "alcohol", 0.8)

### Sugar

findAssocs(corp_dtm, "sugar", 0.7)
```



And the next step is topic modelling  - this allows us to analyse the whole body of bulletins and look for themes or topics - groupings of words within and between documents.

```{r include=FALSE}
library(topicmodels)

corp_lda <- LDA(corp_dtm, k = 6, control = list(seed = 1234))


```


```{r}
corp_lda_tidy <- tidy(corp_lda)
```

```{r, fig.height=12}
corp_lda_tidy %>%
  mutate(row = row_number()) %>%
  group_by(topic) %>%
  top_n(20, beta) %>%
  ungroup() %>%
  arrange(topic, -beta) %>%
  ggplot(aes(reorder(term, row), beta, fill = factor(topic), label = term)) +
  geom_col() +
  geom_text(hjust = 0, size = 10) +
  coord_flip() +
  facet_wrap(~topic, ncol =4) +
  theme_jf() +
  labs(fill = "Topic")
  
  
```


```{r, fig.height=12}
corp_lda_tidy %>%
  group_by(topic) %>%
  top_n(20, beta) %>%
  ungroup() %>%
  arrange(topic, -beta) %>%
  ggplot(aes(term, beta, fill = factor(topic))) +
  geom_point(aes(colour = factor(topic))) +
  geom_line(aes(group = topic)) +
  coord_polar() +
  facet_wrap(~topic, ncol = 4) +
  theme(legend.position = "") +
  theme(axis.text.x = element_text(size = 12))
  
```

### Classifying documents

```{r document classification}
lda_gamma <- tidytext:::tidy.LDA(corp_lda, matrix = "gamma")

lda_gamma

abs_class <- lda_gamma %>%
  group_by(document) %>%
  top_n(1, gamma) %>%
  ungroup() %>%
  arrange(gamma)

abs_class

abs_class %>%
  group_by(topic) %>%
  count()

abs_class %>%
  ggplot(aes(lubridate::ymd(document), as.factor(topic), label = document)) +
  geom_point(aes(colour = topic)) +
  geom_path(aes(colour = topic))+
  geom_text(size = 2, angle = 45, hjust = 0, vjust = 0)+
  coord_cartesian(ylim = c(0, 9)) +
  labs(y = "Topic number" ) +
  theme(legend.position = "")
  


```

