---
title: "Correlation network plots of health profiles"
author: "Julian Flowers"
date: "15/05/2017"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)
```

## Load libraries

```{r load libraries}
library(fingertipsR) ## for interacting with the Fingertips API
library(tidyverse, warn.conflicts = FALSE) ## packages for data manipulation and plotting
library(stringr) ## text manipulation
library(Hmisc) ## data description
library(corrr) ## calcuate correlation matrices
library(ggraph) ## plot as network charts
library(igraph, warn.conflicts = FALSE) ## network tools
```

```{r}

profiles <- profiles()
profiles %>% select(ProfileID, ProfileName) %>% distinct() %>%
  knitr::kable(align = "l")

```



## Select profile and download data

```{r}

options(digits = 2)

profile_name <- filter(profiles, ProfileID == 26) %>% select(ProfileName) %>% distinct()

hp_data <- fingertips_data(ProfileID = 26)

unique(hp_data$IndicatorName) 
```


## Normalise the data

This part of the script:

* converts factor (categorical) variables to character variables
* simplifies column names
* filters the most recent data for each indicator
* creates a unique index for each indicator (indicator+age+sex+period)
* converts a long format table to a cross tab
* removes duplicates
* fills missing data with mean values
* scales (normalises) the dataset (i.e. calculates z-scores for each variable)

```{r}

## scale the data
hp_data1 <- hp_data %>%
  mutate_if(is.factor, as.character) %>%
  filter(CategoryType == "", AreaType == "County & UA" ) %>%
  janitor::clean_names() %>%
  select(indicatorname, areaname, age, sex, timeperiod, value) %>%
  group_by(indicatorname, sex) %>%
  filter(timeperiod == max(timeperiod)) %>%
  mutate(index = paste(sex,  "-",age , "-",timeperiod, "-", indicatorname) ) %>%
  ungroup() %>%
  select(-c(indicatorname, sex, age, timeperiod)) %>%
  distinct() %>%
  spread(index, value) %>%
  janitor::clean_names() %>%
  mutate_if(is.numeric, funs(impute(., mean))) %>%
  mutate_if(is.numeric, funs(scale(.))) 

```

## Create correlation network map

This code:

* Calculates the correlation matrix between all variables
* Extracts well correlated variables (r > 0.7)
* Converts this to graph (network) format
* Plots the correlations as a network map


```{r, fig.height=11, fig.width=8}

## see https://drsimonj.svbtle.com/how-to-create-correlation-network-plots-with-corrr-and-ggraph
hp_cor <- hp_data1 %>%
  select(3:ncol(.)) %>%
  correlate() %>%
  stretch() 

graph_cors <- hp_cor %>% 
  filter(abs(r) > 0.7) %>% 
  graph_from_data_frame(directed = FALSE)

ggraph(graph_cors, layout = "igraph", algorithm = "nicely") +
 geom_edge_link(aes(edge_alpha = abs(r), color = r, label = round(r,2)), label_size = 2) +
  guides(edge_alpha = "none", edge_width = "none") +
  scale_edge_colour_gradientn(limits = c(-1, 1), colors = c("firebrick2", "dodgerblue2")) +
  geom_node_point(color = "black", size = 3) +
  geom_node_point() +
  geom_node_text(aes(label = str_wrap(substring(name, 1,70), 30)), size  = 2, repel = TRUE) +
  theme_graph() +
  labs(title = paste("Correlation map of ", profile_name$ProfileName, "profile"))


```




