---
title: "Smoking test yourself"
author: "Julian Flowers"
date: "10/08/2017"
output: slidy_presentation
---

```{r setup, include=FALSE, cache=TRUE}
knitr::opts_chunk$set(cache = TRUE, echo = FALSE)
```

```{r libraries}
library(shinysense)
library(fingertipsR)
library(tidyverse)
library(stringr)

mydata <- fingertips_data(ProfileID = 19, rank = TRUE, categorytype = TRUE)


```

## Test your knowledge of health data in England

- The following slides show trend in some key indicators of health in England
- The data comes from the Public Health Outcomes Framework
- The trends are incomplete
- Using your mouse, draw what you think the trend is
- The true trend will be revealed
- How did you do?


## Smoking in young people (aged 15)

- Going up or down?

```{r smoking 15,  echo = FALSE}
library(tidyverse)
library(stringr)
library(shinysense)

draw2 <- mydata %>%
  filter(IndicatorID == 91183) %>%
  filter(!is.na(Value)) %>% 
  mutate_if(is.factor, as.character) %>%
  select(AreaName, TimeperiodSortable, Value)

drawx <- draw2 %>%
  mutate(date = as.Date(paste(str_sub(TimeperiodSortable, 1, 4), "01", "01", sep = "-")))

test_data1 <- drawx %>%
  filter(AreaName == "England") %>%
  select(date = date, value = Value)

endDate <- test_data1$date[6]

drawr_widget(
  data = test_data1,
  draw_start = endDate,
  x_key = "date",
  y_key = "value",
  width="100%"
)

```

## Drug deaths

```{r drug deaths}
draw <- mydata %>%
  filter(IndicatorID == 92432, AreaName == "England", Sex == "Persons") %>%
  filter(!is.na(Value), is.na(Category)) %>% 
  mutate_if(is.factor, as.character) %>%
  select(AreaName, TimeperiodSortable, Value)

draw1 <- draw %>%
  mutate(date = as.Date(paste(str_sub(TimeperiodSortable, 1, 4), "01", "01", sep = "-")))


test_data <- draw1 %>%
  filter(AreaName == "England") %>%
  select(date = date, value = Value) 

endDate <- test_data$date[8]

drawr_widget(
  data = test_data,
  draw_start = endDate,
  x_key = "date",
  y_key = "value",
  width="100%"
)
```


## Incidence of TB

```{r incidence of tb}
draw3 <- mydata %>%
  filter(IndicatorID == 91361) %>%
  filter(!is.na(Value), is.na(Category)) %>% 
  mutate_if(is.factor, as.character) %>%
  select(AreaName, TimeperiodSortable, Value)

draw4 <- draw3 %>%
  mutate(date = as.Date(paste(str_sub(TimeperiodSortable, 1, 4), "01", "01", sep = "-")))


test_data2 <- draw4 %>%
  filter(AreaName == "England") %>%
  select(date = date, value = Value)

endDate <- test_data2$date[7]

drawr_widget(
  data = test_data2,
  draw_start = endDate,
  x_key = "date",
  y_key = "value",
  width="100%"
)
```


## Excess winter deaths

```{r winter deaths}
draw5 <- mydata %>%
  filter(IndicatorID == 90360) %>%
  filter(!is.na(Value), is.na(Category), Sex == "Persons") %>% 
  mutate_if(is.factor, as.character) %>%
  select(AreaName, TimeperiodSortable, Value)

draw6 <- draw5 %>%
  mutate(date = as.Date(paste(str_sub(TimeperiodSortable, 1,4), "01", "01", sep = "-")))


test_data3 <- draw6 %>%
  filter(AreaName == "England") %>%
  select(date = date, value = Value)

endDate <- test_data3$date[7]

drawr_widget(
  data = test_data3,
  draw_start = endDate,
  x_key = "date",
  y_key = "value",
  width="100%"
)
```


## New

```{r}
draw3 <- mydata %>%
  filter(IndicatorID == 91361) %>%
  filter(!is.na(Value), is.na(Category)) %>% 
  mutate_if(is.factor, as.character) %>%
  select(AreaName, TimeperiodSortable, Value)

draw4 <- draw3 %>%
  mutate(date = as.Date(paste(str_sub(TimeperiodSortable, 1, 4), "01", "01", sep = "-")))


test_data2 <- draw4 %>%
  filter(AreaName == "England") %>%
  select(date = date, value = Value)

endDate <- test_data2$date[7]

drawr_widget(
  data = test_data2,
  draw_start = endDate,
  x_key = "date",
  y_key = "value",
  width="100%"
)
```


## Incidence of TB

```{r }
draw3 <- mydata %>%
  filter(IndicatorID == 91361) %>%
  filter(!is.na(Value), is.na(Category)) %>% 
  mutate_if(is.factor, as.character) %>%
  select(AreaName, TimeperiodSortable, Value)

draw4 <- draw3 %>%
  mutate(date = as.Date(paste(str_sub(TimeperiodSortable, 1, 4), "01", "01", sep = "-")))


test_data2 <- draw4 %>%
  filter(AreaName == "England") %>%
  select(date = date, value = Value)

endDate <- test_data2$date[7]

drawr_widget(
  data = test_data2,
  draw_start = endDate,
  x_key = "date",
  y_key = "value",
  width="100%"
)
```


